Baseline centering problem

### 2018-11-10 paper/advup_uniquesol.tex

\paragraph{A variant to make the learning of $A$ and $b$ insensitive to
translations of $V$, and the dynamics independent from
$\alpha_3/\alpha_5$.} Consider the modified off-policy Bellman equation
for all states $s$,
\begin{equation}
0=\E \left(
r(s,a)-b+\gamma V(s')-V(s)+(1-\gamma)V(\tilde s)
\right)
\end{equation}
where $a$ is from some policy $\pi$, $s'$ is the next state given $s$ and
$a$, and $\tilde s$ is a random state taken from any empirical
distribution $\rho$.

@@

Let us minimize this by TD with $s$ taken from the same
empirical distribution $\rho$, adding a gradient term to minimize
$\frac{\alpha}{2} (\E_\rho V)^2$ as before.

The TD step is
\begin{equation}
\E_s \,\partial_\theta V(s)\left(
r(s,a)-b+\gamma V(s')-V(s)+(1-\gamma)V(\tilde s)-\alpha V(\tilde s)
\right)
\end{equation}
and its fixed point solves $r(s,a)-b+\gamma V(s')-V(s)+(1-\gamma)V(\tilde
s)-\alpha V(\tilde s)$ for all $s$. The unique solution for $V$ given $b$ is
\begin{equation}
V(s)=V^\pi(s)-\E_\rho V^\pi-\frac{b-(1-\gamma)\E_\rho V^\pi}{\alpha}
\end{equation}

Meanwhile, $b$ is trained to minimize $\E_s\E_{s'|s,a}\E_{\tilde s} (r-b+\gamma
V(s')-V(s)+(1-\gamma)V(\tilde s))^2$. Given $V$, the unique solution is
\begin{equation}
b=\E_s\E_{s'|s,a}\E_{\tilde s}(r+\gamma
V(s')-V(s)+(1-\gamma)V(\tilde s))
\end{equation}
Substituting the equilibrium value for $V$, $V=V^\pi+\mathrm{cst}$ yields
\begin{equation}
b=(1-\gamma)\E_{\tilde s} V^\pi(\tilde s)=(1-\gamma)\E_\rho V^\pi
\end{equation}
whatever $\mathrm{cst}$ is. Thus, for this optimal value of $b$, we have
\begin{equation}
V(s)=V^\pi(s)-\E_\rho V^\pi
\end{equation}

>>>>

\paragraph{A variant to make the learning of $A$ and $b$ insensitive to
translations of $V$, and the dynamics independent from
$\alpha_3/\alpha_5$.} Consider the modified off-policy Bellman equation
for all states $s$,
\begin{equation}
0=\E \left(
r(s,a)-b+\gamma V(s')-V(s)+(1-\gamma)V(\tilde s)
\right)
\end{equation}
where $a$ is from some policy $\pi$, $s'$ is the next state given $s$ and
$a$, and $\tilde s$ is a random state, independent from $s$, taken from any empirical
distribution $\tilde \rho$.

@@

Let us minimize this by TD with $s$ taken from some other empirical
distribution $\rho$, and add a TD-style gradient term to minimize
$\frac{\alpha}{2} (\E_{\tilde \rho} V)^2$. Interestingly we do not need
to have $\rho=\tilde \rho$: in that case, we are not taking a true
gradient of $\frac{\alpha}{2} (\E_{\tilde \rho} V)^2$ but a TD-style
gradient, but it still has the right fixed point, as we show now.

The TD step is
\begin{equation}
\E_s \,\partial_\theta V(s)\left(
r(s,a)-b+\gamma V(s')-V(s)+(1-\gamma)V(\tilde s)-\alpha V(\tilde s)
\right)
\end{equation}
and its fixed point solves $0=\E[r(s,a)-b+\gamma V(s')-V(s)+(1-\gamma)V(\tilde
s)-\alpha V(\tilde s)]$ for all $s$.\footnote{Here I assume the family
$V_\theta$ is
overparameterized: in that case, if the average TD step is $0$ then the
right-hand side vanishes for every $s$. If not, the difference between $V(s)$ and the solution
presented is orthogonal to the span of $\partial_\theta V$ wrt some weird
metric.} The unique solution for $V$ given $b$ is
\begin{equation}
V(s)=V^\pi(s)-\E_{\tilde \rho} V^\pi-\frac{b-(1-\gamma)\E_{\tilde \rho} V^\pi}{\alpha}
\end{equation}

Meanwhile, $b$ is trained to minimize $\E_s\E_{s'|s,a}\E_{\tilde s} (r-b+\gamma
V(s')-V(s)+(1-\gamma)V(\tilde s))^2$. Given $V$, the unique solution is
\begin{equation}
b=\E_s\E_{s'|s,a}\E_{\tilde s}(r+\gamma
V(s')-V(s)+(1-\gamma)V(\tilde s))
\end{equation}
Substituting the equilibrium value for $V$, $V=V^\pi+\mathrm{cst}$ yields
\begin{equation}
b=(1-\gamma)\E_{\tilde s} V^\pi(\tilde s)=(1-\gamma)\E_{\tilde \rho} V^\pi
\end{equation}
whatever $\mathrm{cst}$ is. Thus, for this optimal value of $b$, we have
\begin{equation}
V(s)=V^\pi(s)-\E_{\tilde \rho} V^\pi
\end{equation}

Notably the distributions of $s$ and $\tilde s$ do not have to match. We may even take a single value $s_0$ for $\tilde s$ as a reference point
($\tilde \rho$ a Dirac mass at $s_0$): $V$ would converge to
$V^\pi-V^\pi(s_0)$, and $b$ to $(1-\gamma)V^\pi(s_0)$. This may be the
simplest practical solution. (If changing $s_0$ from time to time, $V$
would have to be shifted by the change in $V^\pi(s_0)$.)
